Wigglyhedra

### 2024-07-12 draft/wigglyhedra.tex
maximise front page picture size
Several updates (some in progress).

\centerline{\includegraphics[scale=1.6]{wigglyFlipGraph}}

@@

\begin{remark}
\label{rem:numberWigglyPseudotriangulations}
The number~$wp_n$ of wiggly pseudotriangulations of~$n+2$ points is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly pseudotriangulations of~$n+2$ points with~$i$ \defn{final} internal wiggly arcs (\ie ending at the last point~$n+1$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPseudotriangulations}.
We use this additional variable~$x$ to obtain a recursive formula for~$wp_n(x)$.
%Indeed, any wiggly pseudotriangulation on~$n+1$ points can be obtained from a wiggly pseudotriangulation on~$n+2$ points by collapsing the points~$n$ and~$n+1$.
Indeed, any wiggly pseudotriangulation of~$n+2$ points can be obtained from a wiggly pseudotriangulation of~$n+1$ points by inserting its wiggly pseudotriangle with a hinge at~$n$.
%There are two kinds of wiggly pseudotriangulations of~$n+2$ points:
%\begin{itemize}
%\item those with the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the position of the only corner of~$t_n$ which is not at~$n+1$,
%\item those without the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the positions of the two corners of~$t_n$ which are not at~$n+1$.
%\end{itemize}
Each wiggly pseudotriangulation of~$n+1$ points with $i$ final internal wiggly arcs contributes to
\begin{itemize}
\item $i+2$ wiggly pseudotriangulations of~$n+2$ points with $i+2$ final internal wiggly arcs, containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,
\item $j+1$ wiggly pseudotriangulations of~$n+2$ points with $j$ final internal wiggly arcs, not containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,  for all~$0 \le j \le i+1$.
\end{itemize}
\pagebreak
This translates to the recursive formula %for all~$n \ge 1$,
\begin{align*}
wp_{n}(x) & = \overbracket[.5pt]{x \cdot \frac{\partial}{\partial x} \big( x^2 wp_{n-1}(x) \big)}^{\substack{\text{wiggly pseudotriangulations} \\ \text{containing } (n, n+1, \varnothing, \varnothing)}} + \overbracket[.5pt]{\frac{\partial}{\partial x}\Big( \smash{\frac{wp_{n-1}(1)-x^3wp_{n-1}(x)}{1-x}} \Big)}^{\substack{\qquad\; \text{wiggly pseudotriangulations} \qquad\;  \\ \text{not containing } (n, n+1, \varnothing, \varnothing)}} \\
& = \frac{1}{(1 - x)^2} \big( wp_{n-1}(1) + x^2 (2 x^2 - 2 x - 1) wp_{n-1}(x) + x^4 (x - 1) wp_{n-1}’(x) \big).
\end{align*}
This enables us to quickly compute~$wp_n(x)$ and we obtain the numbers of wiggly pseudotriangulations~$wp_n$ by evaluating~$wp_n(x)$ at~$x = 1$.
%This enables us to quickly compute the generating series of the number of wiggly pseudotriangulations
%\[
%\sum_{n \ge 0} wp_n(1) y^n = 1 + 2 y + 14 y^2 + 176 y^3 + 3232 y^4 + 78384 y^5 + 2366248 y^6 + 85534176 y^7 + 3602770400 y^8 + 
%% 173300710720 y^9 + 9373542317760 y^10 + 563142033172480 y^11 + 37206559614499840 y^12 + 2681213937595142656 y^13 +
%\dots
%\]
\end{remark}

@@

There are various possible proofs for the connectiveness.
For instance, it will follow easily from \cref{prop:bijection}.
To avoid to rely on this forward reference, we sketch an alternative direct proof by induction on~$n$.
We claim that if a wiggly pseudotriangulation has some final internal wiggly arcs (\ie ending at the last point~$n+1$), then at least one of them can be flipped to a non-final wiggly arc.
This implies that any wiggly pseudotriangulation can be transformed by flips to a wiggly pseudotriangulation with no final internal wiggly arc.
All wiggly pseudotriangulations with no final internal wiggly arc contain the arcs~$(0,n,\varnothing,[n-1])$ and~$(0,n,[n-1],\varnothing)$, since the latter are compatible with all internal wiggly arcs but the final ones.
Hence, they induce a subgraph of~$\wigglyFlipGraph_n$ isomorphic to~$\wigglyFlipGraph_{n-1}$, which is connected by induction.
We conclude that~$\wigglyFlipGraph_n$ is itself connected.

To see the claim, let~$i$ be maximal such that~$T$ contains two distinct wiggly arcs starting at~$i$ and ending at~$n+1$ (it is well-defined as we have the external wiggly arcs~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$ for~$i = 0$).
Consider a final internal wiggly arc~$\alpha$ of~$T$ starting weakly after~$i$, and let~$\alpha'$ be the wiggly arc obtained by flipping~$\alpha$ in~$T$.
If~$\alpha'$ is final, then at least one pseudotriangle~$t$ of~$T$ incident to~$\alpha$ has two corners at~$n+1$.
Let~$h$ be the hinge of~$t$ and~$\beta, \beta'$ be the two wiggly arcs of~$t$ joining~$h$ to~$n+1$.
As~$i < h \le n$ the maximality of~$i$ imposes that~$\beta = \beta'$.
Flipping $\beta = \beta'$ in~$T$ thus yields a wiggly arc ending at~$h \le n$, hence a non-final arc.
\end{proof}

We will even consider an oriented version of the wiggly flip graph~$\wigglyFlipGraph_n$.
The following definition is illustrated in \cref{fig:incompatible2}.

@@

\pagebreak
We will also need the following observation about exchangeable arcs, illustrated in \cref{fig:incompatible3}.

@@

\begin{remark}
\label{rem:wigglyComplexAutomorphism}
The vertical and horizontal reflections provide two obvious automorphisms of the wiggly complex~$\wigglyComplex_n$ (and thus of the wiggly flip graph~$\wigglyFlipGraph_n$).
Moreover, the vertical reflection is an anti-automorphism of the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$ (while the horizontal reflection is not).
\end{remark}

@@

\begin{remark}
\label{rem:numberWigglyPermutations}
The number~$wp_n$ of wiggly permutations of~$[2n]$ is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly permutations of~$[2n]$ with $i$ \defn{admissible gaps} (\ie gaps~$\gamma$ between two consecutive positions such that there is no~$j \in [n]$ such that the value~$2j$ appears before the gap~$\gamma$ while the value~$2j-1$ appears after the gap~$\gamma$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPermutations}.
Any wiggly permutation of~$[2n]$ can be obtained from a wiggly permutation of~$[2n-2]$ by inserting the last values~$2n-1$ and~$2n$.
Each wiggly permutation of~$[2n-2]$ with $i$ admissible gaps contributes to
\begin{itemize}
\item $i+2$ wiggly permutations of~$[2n]$ with $i+2$ admissible gaps and where~$2n-1$ appears before~$2n$ (they must appear consecutively),
\item $j+1$ wiggly permutations of~$[2n]$ with $j$ admissible arcs and where $2n$ appears before~$2n-1$, for all~$0 \le j \le i+1$.
\end{itemize}
We thus obtain the same recursive equation on~$wp_n(x)$ as in \cref{rem:numberWigglyPseudotriangulations}.
\end{remark}

@@

The aim of this section is to explain a categorical interpretation of wiggly arcs.
We first present a particular triangulated category (\cref{subsec:category}), and recall results of Khovanov--Seidel~\cite{kho.sei:02} (\cref{subsec:KSCurveModel}) that interpret the objects of this category as wiggly arcs, such that intersections between wiggly arcs correspond to morphisms in this category.
We then explain how wiggly pseudotriangulations govern the combinatorial structure of a natural categorical decomposition (\cref{subsec:cohomologyFiltrations}).
We also discuss categorical conditions for two wiggly arcs to be compatible as in~\cref{def:compatible} (\cref{subsec:categoricalCompatibility}).

@@

We begin by describing the category of interest, which is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra~\(A_m\), called the \defn{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let~\(e_i\) denote the idempotent path at the vertex \(i\), and let \((i, j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by~\((i_1, i_2, \dots, i_n)\).

@@

We now describe the \(t\)-structure induced by the full subcategory of \defn{linear complexes} of \(\mathcal{C}_m\).
A variant of this construction is also described in~\cite[Sect.~2.5]{lic.que:21}.
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\) as a bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[
P^{\bullet} = \bigoplus_{g,h,i} M_i\langle g \rangle[h],
\]
where \(g\) is the internal grading shift, and \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).
Set \(P^h\) to be the direct sum of all summands that have homological grading shift \(h\).
That is, in the notation above, we have
\[
P^{h} = \bigoplus_{g,i} M_i\langle g \rangle[h].
\]

@@

The next statement, easy to check from the definitions, is mentioned in~\cite[Sect.~3.3]{bayer2011tour}.

@@

\begin{proposition}
\label{prop:cohomology-pieces}
Let \(\alpha\) be any curve, and let \(X = X^{\bullet}\) be a chain complex that corresponds to \(\alpha\) under the Khovanov--Seidel reconstruction.
Recall that for any integer \(j\), the cohomology object \(H^j(X)\) is an object of \(\mathcal{A}_m\), and hence a direct sum of indecomposable objects of \(\mathcal{A}_m\).
\begin{enumerate}
\item The object \(X\) is an indecomposable spherical object of \(\mathcal{A}_m\) up to shift if and only if \(\alpha\) is a wiggly arc. In this case, \(H^j(X) \neq 0\) for exactly one \(j\), such that \(X \cong H^j(X)[-j]\).
\item If \(\alpha\) is a general curve, each indecomposable summand of \(H^j(X)\) for any \(j\) is a spherical indecomposable object of \(\mathcal{A}_m\) that arises from a wiggly arc, up to shift.
\end{enumerate}
\end{proposition}

\begin{proof}
  Let \(X\) be a complex built from an arbitrary (not necessarily wiggly) curve \(\alpha\).
  The results of~\cite{kho.sei:02} imply that \(X\) is spherical and indecomposable.
  Let us now analyse the indecomposable summands of its cohomology objects.

  Recall that the procedure to build up \(X\) from \(\alpha\) explicitly describes the differentials.    
  Each component of these differentials is a map of the form
  \[P_i\langle g \rangle[h] \to P_j\langle g' \rangle[h-1].\]
  Let us see how these maps interact with the level.

  First consider a non-zero component \(P_i\langle g \rangle[h] \to P_j\langle g' \rangle[h-1]\) for \(i \neq j\).
  We see from~\cref{rem:generating-morphisms} that \(g' = g+1\).
  In other words, this component has the form
  \(P_i\langle g \rangle[h] \to P_j\langle g+1 \rangle[h-1]\),
  and hence does not change the level.

  Now consider a non-zero component \(P_i\langle g \rangle[h] \to P_i\langle g' \rangle[h-1]\).
  Once again, it is clear from~\cref{rem:generating-morphisms} that we either have \(g' = g\) or \(g' = g + 2\).
  Any component of the form \(P_i \langle g \rangle[h] \to P_i\langle g \rangle[h-1]\)
  is necessarily a multiple of the identity map, which do not appear in the Khovanov--Seidel complex.
  The only remaining possibility is a ``loop'' map \(P_i \langle g \rangle[h] \to P_i\langle g+2 \rangle[h-1]\), which raises the level by one.

  Computing the \(j\)th cohomology \(H^j(X)\) simply amounts to truncating the complex to level \(j\).
  Explicitly, the truncation is a new chain complex in which we only keep the terms of \(X\) that have level \(j\), together with all maps between them, and forget all the other terms.
  From the calculation above, we see that the resulting complex only has components in its differential of the form \(P_i\langle g \rangle[h] \to P_j\langle g+1 \rangle[h-1]\) for \(i \neq j\); in particular, we kill all the ``loop'' components.
  The resulting chain complex breaks up into indecomposable summands, each of which is an object of~\(\mathcal{A}_m[j]\), and this direct sum is \(H^j(X)[j]\).

  The object \(X\) is in the heart up to shift if and only if it is concentrated in a single cohomological degree \(j\).
  Truncating to level \(j\) as described above, we do not delete any of the terms of \(X\).
  This can happen if and only if the differential of \(X\) contains no loop components,
  which appear in the complex if and only if \(\alpha\) loops around a point.
  Thus \(X\) is in the heart up to shift if and only if \(\alpha\) is  \(x\)-monotone.
  This proves the first part of the proposition.

  More generally, consider the indecomposable pieces of the level-\(j\) truncations of \(X\) for various~\(j\), up to shift.
  As explained above, in terms of the curve \(\alpha\), this corresponds to breaking up the curve~\(\alpha\) into \(x\)-monotone portions.
  More precisely, we begin at an endpoint \(p_0\) of \(\alpha\) and follow the curve until it either ends at a point \(p_1\), or stops being \(x\)-monotone by bending around a point \(p_1\).
  In this case, we take the portion of \(\alpha\) between \(p_0\) and \(p_1\), calling it \(\alpha_1\), and setting its endpoints to be \(p_0\) and \(p_1\) respectively.
  Note that \(p_0 \neq p_1\) because \(\alpha\) is \(x\)-monotone between \(p_0\) and \(p_1\).
  We continue with the remainder of \(\alpha\) starting at \(p_1\), greedily taking the maximal \(x\)-monotone portion each time, to obtain pieces \(\alpha_1,\alpha_2, \ldots, \alpha_k\), which are all wiggly arcs.
  These wiggly arcs correspond to complexes \(X_1, \ldots, X_k\), which are indecomposable spherical objects of \(\mathcal{A}_m\), up to shift.
  It is clear from the construction of the associated complex that these objects, up to shift, are exactly the indecomposable pieces of the cohomology truncations \(H^j(X)\) for various \(j\).
  This argument proves the second part of the proposition.
\end{proof}

@@

  Recall that \(\beta\) and \(\beta'\) are two pieces of the curve \(\alpha\).
  Therefore \(\beta\) and \(\beta'\) cannot cross, because~\(\alpha\) does not cross itself.
  Suppose that \(\beta\) and \(\beta'\) are non-pointed, and without loss of generality, suppose that \(i' = j\).
  This means that the piece \(\beta\) was cut from \(\alpha\) either because \(i' = j\) is an endpoint of \(\alpha\), or because \(\alpha\) loops around the right of the point \(i'\).
  Moreover, if \(i'\) is an endpoint of \(\alpha\), then the portion of \(\alpha\) closest to \(i'\) lies to the left of \(i'\).

>>>>

\centerline{\includegraphics[scale=1.75]{wigglyFlipGraph}}

@@

\begin{remark}
\label{rem:numberWigglyPseudotriangulations}
The number~$wp_n$ of wiggly pseudotriangulations of~$n+2$ points is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly pseudotriangulations of~$n+2$ points with~$i$ \defn{final} internal wiggly arcs (\ie ending at the last point~$n+1$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPseudotriangulations}.
We use this additional variable~$x$ to obtain a recursive formula for~$wp_n(x)$.
%Indeed, any wiggly pseudotriangulation on~$n+1$ points can be obtained from a wiggly pseudotriangulation on~$n+2$ points by collapsing the points~$n$ and~$n+1$.
Indeed, any wiggly pseudotriangulation of~$n+2$ points can be obtained from a wiggly pseudotriangulation of~$n+1$ points by inserting a single wiggly pseudotriangle with a hinge at~$n$.
%There are two kinds of wiggly pseudotriangulations of~$n+2$ points:
%\begin{itemize}
%\item those with the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the position of the only corner of~$t_n$ which is not at~$n+1$,
%\item those without the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the positions of the two corners of~$t_n$ which are not at~$n+1$.
%\end{itemize}
Each wiggly pseudotriangulation of~$n+1$ points with $i$ final internal wiggly arcs contributes to:
\begin{itemize}
\item $i+2$ wiggly pseudotriangulations of~$n+2$ points with $i+2$ final internal wiggly arcs, containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,
\item $j+1$ wiggly pseudotriangulations of~$n+2$ points with $j$ final internal wiggly arcs, not containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,  for all~$0 \le j \le i+1$.
\end{itemize}
\pagebreak
This translates to the recursive formula %for all~$n \ge 1$,
\begin{align*}
wp_{n}(x) & = \overbracket[.5pt]{x \cdot \frac{\partial}{\partial x} \big( x^2 wp_{n-1}(x) \big)}^{\substack{\text{wiggly pseudotriangulations} \\ \text{containing } (n, n+1, \varnothing, \varnothing)}} + \overbracket[.5pt]{\frac{\partial}{\partial x}\Big( \smash{\frac{wp_{n-1}(1)-x^3wp_{n-1}(x)}{1-x}} \Big)}^{\substack{\qquad\; \text{wiggly pseudotriangulations} \qquad\;  \\ \text{not containing } (n, n+1, \varnothing, \varnothing)}} \\
& = \frac{1}{(1 - x)^2} \big( wp_{n-1}(1) + x^2 (2 x^2 - 2 x - 1) wp_{n-1}(x) + x^4 (x - 1) wp_{n-1}’(x) \big).
\end{align*}
This enables us to quickly compute~$wp_n(x)$ and we obtain the numbers of wiggly pseudotriangulations~$wp_n$ by evaluating~$wp_n(x)$ at~$x = 1$.
%This enables us to quickly compute the generating series of the number of wiggly pseudotriangulations
%\[
%\sum_{n \ge 0} wp_n(1) y^n = 1 + 2 y + 14 y^2 + 176 y^3 + 3232 y^4 + 78384 y^5 + 2366248 y^6 + 85534176 y^7 + 3602770400 y^8 + 
%% 173300710720 y^9 + 9373542317760 y^10 + 563142033172480 y^11 + 37206559614499840 y^12 + 2681213937595142656 y^13 +
%\dots
%\]
\end{remark}

@@

There are various possible proofs for the connectedness.
For instance, it will follow easily from \cref{prop:bijection}.
To avoid relying on this forward reference, we sketch an alternative direct proof by induction on~$n$.
We claim that if a wiggly pseudotriangulation has a non-zero number of final internal wiggly arcs (\ie ending at the last point~$n+1$), then at least one of them can be flipped to a non-final wiggly arc.
This implies that any wiggly pseudotriangulation can be transformed by flips to a wiggly pseudotriangulation with no final internal wiggly arc.
All wiggly pseudotriangulations with no final internal wiggly arc contain the arcs~$(0,n,\varnothing,[n-1])$ and~$(0,n,[n-1],\varnothing)$, since the latter are compatible with all internal wiggly arcs but the final ones.
Hence, they induce a subgraph of~$\wigglyFlipGraph_n$ isomorphic to~$\wigglyFlipGraph_{n-1}$, which is connected by induction.
We conclude that~$\wigglyFlipGraph_n$ is itself connected.

To see the claim, let~$i$ be maximal such that~$T$ contains two distinct wiggly arcs starting at~$i$ and ending at~$n+1$ (it is well-defined as we have the external wiggly arcs~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$ for~$i = 0$).
Consider a final internal wiggly arc~$\alpha$ of~$T$ starting weakly after~$i$, and let~$\alpha'$ be the wiggly arc obtained by flipping~$\alpha$ in~$T$.
If~$\alpha'$ is final, then by~\cref{prop:diagonalsPseudoquadrangle}, at least one pseudotriangle~$t$ of~$T$ incident to~$\alpha$ has two corners at~$n+1$.
Let~$h$ be the hinge of~$t$ and~$\beta, \beta'$ be the two wiggly arcs of~$t$ joining~$h$ to~$n+1$.
As~$i < h \le n$ the maximality of~$i$ imposes that~$\beta = \beta'$.
Flipping $\beta = \beta'$ in~$T$ thus yields a wiggly arc ending at~$h \le n$, hence a non-final arc.
\end{proof}

We will also consider an oriented version of the wiggly flip graph~$\wigglyFlipGraph_n$.
The following definition is illustrated in \cref{fig:incompatible2}.

@@

\pagebreak
We will need the following observation about exchangeable arcs, illustrated in \cref{fig:incompatible3}.

@@

\begin{remark}
\label{rem:wigglyComplexAutomorphism}
The vertical and horizontal reflections provide two obvious automorphisms of the wiggly complex~$\wigglyComplex_n$ (and thus of the wiggly flip graph~$\wigglyFlipGraph_n$).
Moreover, the vertical reflection is an anti-automorphism of the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$ (while the horizontal reflection is not).\asilata{Check, is the horizontal reflection nothing at all? Is it actually an automorphism?}
\end{remark}

@@

\begin{remark}
\label{rem:numberWigglyPermutations}
The number~$wp_n$ of wiggly permutations of~$[2n]$ is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly permutations of~$[2n]$ with $i$ \defn{admissible gaps} (\ie gaps~$\gamma$ between two consecutive positions such that there is no~$j \in [n]$ such that the value~$2j$ appears before the gap~$\gamma$ while the value~$2j-1$ appears after the gap~$\gamma$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPermutations}.
Any wiggly permutation of~$[2n]$ can be obtained from a wiggly permutation of~$[2n-2]$ by inserting the last values~$2n-1$ and~$2n$.
Each wiggly permutation of~$[2n-2]$ with $i$ admissible gaps contributes to:
\begin{itemize}
\item $i+2$ wiggly permutations of~$[2n]$ with $i+2$ admissible gaps and where~$2n-1$ appears before~$2n$ (they must appear consecutively),
\item $j+1$ wiggly permutations of~$[2n]$ with $j$ admissible arcs and where $2n$ appears before~$2n-1$, for all~$0 \le j \le i+1$.
\end{itemize}
We thus obtain the same recursive equation on~$wp_n(x)$ as in \cref{rem:numberWigglyPseudotriangulations}.
\end{remark}

@@

The aim of this section is to explain a categorical interpretation of wiggly arcs.
We first present a particular triangulated category (\cref{subsec:category}), and recall results of Khovanov--Seidel~\cite{kho.sei:02} (\cref{subsec:KSCurveModel}) that interpret the objects of this category as wiggly arcs, such that intersections between wiggly arcs correspond to morphisms in this category.
We then explain how wiggly pseudotriangulations govern the combinatorial structure of a natural categorical decomposition of objects (\cref{subsec:cohomologyFiltrations}).
We also discuss categorical conditions for two wiggly arcs to be compatible as in~\cref{def:compatible} (\cref{subsec:categoricalCompatibility}).

@@

We begin by describing the category of interest, which is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of finitely generated \(A_m\)-modules.
The algebra~\(A_m\), called the \defn{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let~\(e_i\) denote the idempotent path at the vertex \(i\), and let \((i, j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by~\((i_1, i_2, \dots, i_n)\).

@@

We now describe the \(t\)-structure induced by the full subcategory of \defn{linear complexes} of \(\mathcal{C}_m\).
A variant of this construction is also described in~\cite[Sect.~2.5]{lic.que:21}.
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\) as a bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[
P^{\bullet} = \bigoplus_{g,h,i} M_i\langle g \rangle[h],
\]
where \(g\) is the internal grading shift, \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).
Set \(P^h\) to be the direct sum of all summands that have homological grading shift \(h\).
That is, in the notation above, we have
\[
P^{h} = \bigoplus_{g,i} M_i\langle g \rangle[h].
\]

@@

The next statement, easy to check from the definitions, is mentioned \eg in~\cite[Sect.~3.3]{bayer2011tour}.

@@

Any object \(X\) obtained via the Khovanov--Seidel correspondence is~\defn{spherical}; that is, \(\Hom^{g,0}(X,X)\) is \(k\) when \(g = 0,2\), and \(\Hom^{g,h}(X,X) = 0\) in all other cases.
A spherical object \(X\) is necessarily indecomposable as \(\Hom^{0,0}(X,X) = k\).

\begin{proposition}
\label{prop:cohomology-pieces}
Let \(\alpha\) be any curve, and let \(X = X^{\bullet}\) be a chain complex that corresponds to \(\alpha\) under the Khovanov--Seidel reconstruction.
Recall that for any integer \(j\), the cohomology object \(H^j(X)\) is an object of \(\mathcal{A}_m\), and hence a direct sum of indecomposable objects of \(\mathcal{A}_m\).
\begin{enumerate}
\item The object \(X\) is a (spherical) object of \(\mathcal{A}_m\) up to shift if and only if \(\alpha\) is a wiggly arc. In this case, \(H^j(X) \neq 0\) for exactly one \(j\), such that \(X \cong H^j(X)[-j]\).
\item If \(\alpha\) is a general curve, each indecomposable summand of \(H^j(X)\) for any \(j\) is a spherical object of \(\mathcal{A}_m\) that arises from a wiggly arc, up to shift.
\end{enumerate}
\end{proposition}

\begin{proof}
  Let \(X\) be a complex built from an arbitrary (not necessarily wiggly) curve \(\alpha\).
  Recall from~\cite{kho.sei:02} that \(X\) is spherical, and hence indecomposable.
  Let us now analyse the indecomposable summands of its cohomology objects.

  We recall that the procedure to build up \(X\) from \(\alpha\) explicitly describes the differential.
  We analyse the possible non-zero components of the differential, checking how each interacts with the level.
  The only two possibilities are as follows:
  \begin{enumerate}
  \item a map \(P_i\langle g \rangle[h] \to P_{i\pm1}\langle g+1 \rangle[h-1]\), namely multiplication on the right by the path \((i,i\pm1)\), which does not change the level; or
  \item a map \(P_i\langle g \rangle[h] \to P_{i}\langle g+2 \rangle[h-1]\), namely multiplication on the right by the ``loop'' path \((i,i\pm1, i)\), which raises the level by one.
  \end{enumerate}
  The first possibility appears each time the curve wiggles either above the \(i\)th point or below the \(i-1\)th point.
  The second possibility appears each time the curve makes a \(U\)-turn around the \(i\)th or \(i-1\)th point.

  Computing the \(j\)th cohomology \(H^j(X)\) simply amounts to truncating the complex to level \(j\).
  Explicitly, the truncation is a new chain complex in which we only keep the terms of \(X\) that have level \(j\), together with all maps between them, and forget all the other terms.
  From the calculation above, we see that the resulting complex only has components in its differential of the first type, namely \(P_i\langle g \rangle[h] \to P_{i\pm1}\langle g+1 \rangle[h-1]\) where \(g + h = j\).
  We kill all components of the second type (the ``loop'' components).
  The resulting chain complex breaks up into indecomposable summands, each of which is an object of~\(\mathcal{A}_m[j]\), and this direct sum is \(H^j(X)[j]\).

  The object \(X\) is in the heart up to shift if and only if it is concentrated in a single level \(j\).
  Truncating to level \(j\) as described above, we do not delete any of the terms of \(X\).
  This can happen if and only if the differential of \(X\) contains no loop components,
  which appear in the complex if and only if \(\alpha\) makes a U-turn around one of the \(n+2\) points.
  Thus \(X\) is in the heart up to shift if and only if \(\alpha\) is  \(x\)-monotone.
  This proves the first part of the proposition.

  More generally, consider the indecomposable pieces of the level-\(j\) truncations of \(X\) for various~\(j\), up to shift.
  As explained above, in terms of the curve \(\alpha\), this corresponds to breaking up the curve~\(\alpha\) into \(x\)-monotone portions.
  More precisely, we begin at an endpoint \(p_0\) of \(\alpha\) and follow the curve until it either ends at a point \(p_1\), or stops being \(x\)-monotone by bending around a point \(p_1\).
  In this case, we take the portion of \(\alpha\) between \(p_0\) and \(p_1\), calling it \(\alpha_1\), and setting its endpoints to be \(p_0\) and \(p_1\) respectively.
  Note that \(p_0 \neq p_1\) because \(\alpha\) is \(x\)-monotone between \(p_0\) and \(p_1\).
  We continue with the remainder of \(\alpha\) starting at \(p_1\), greedily taking the maximal \(x\)-monotone portion each time, to obtain pieces \(\alpha_1,\alpha_2, \ldots, \alpha_k\), which are all wiggly arcs.
  These wiggly arcs correspond to complexes \(X_1, \ldots, X_k\), which are spherical objects of \(\mathcal{A}_m\), up to shift.
  It is clear from the construction of the associated complex that these objects, up to shift, are exactly the indecomposable pieces of the cohomology truncations \(H^j(X)\) for various \(j\).
  This argument proves the second part of the proposition.
\end{proof}

@@

  Recall that \(\beta\) and \(\beta'\) are two pieces of the curve \(\alpha\).
  Therefore \(\beta\) and \(\beta'\) cannot cross, because~\(\alpha\) does not cross itself.
  Suppose that \(\beta\) and \(\beta'\) are non pointed, and without loss of generality, suppose that \(i' = j\).
  This means that the piece \(\beta\) was cut from \(\alpha\) either because \(i' = j\) is an endpoint of \(\alpha\), or because \(\alpha\) loops around the right of the point \(i'\).
  Moreover, if \(i'\) is an endpoint of \(\alpha\), then the portion of \(\alpha\) closest to \(i'\) lies to the left of \(i'\).
